---
title: "Switching Guidelines"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{switching-guidelines}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Purpose

Using `ideacolors` with `ggplot2` allows you to add IDEA's or Camp RIO's color palettes with familiar `ggplot2` syntax. These branding guidelines change occasionally, and your visualizations will need to be adapated to use the updated guidelines. However, if you are trying to reproduce a visualization from an older branding guideline for consistency, then you'll need to override **the default behavior of `ideacolors` which uses the newest branding guidelines available**.

### Branding guidelines available

Below are the color palettes available from current and past branding guidelines:

- IDEA: **2024 (default)**, 2019
- Camp RIO: **2022 (default)**

## How it works

Suppose you have a visualization that uses an accent color, say `idea_colors_2019$magenta` or `idea_colors_2024$vermillion`, and you want to keep your visualizations consistent between different reports. To specify the branding guidelines for that visualization, you may need to use the `year` argument.

### Option 1: use the most recent guidelines available

If you do not need to use prior branding guidelines, common functions like `scale_color_idea()`, `scale_fill_idea()`, `scale_color_camp_rio()` and `scale_fill_camp_rio()` will default to the most recent guidelines available (in this case, 2024).

```{r use_2024}
library(ggplot2)
library(ideacolors)

# uses most recent branding guidelines
cars <- ggplot(mtcars) +
  geom_histogram(aes(x = mpg,
                     fill = as.factor(cyl))) +
  scale_fill_idea()
  
cars
```


### Option 2: explicitly specify branding guidelines

If you want to specify explicitly which guidelines you will use, then add the `year` argument to the `scale_*` functions. For example, using the 2019 guidelines would produce this graph:

```{r use_2019}
cars +
  scale_fill_idea(year = 2019)
```

Note that using `scale_fill_idea(year = 2024)` would produce the same graph using `scale_fill_idea()`, since those are the most recent guidelines. Similarly, using `scale_fill_camp_rio(year = 2022)` would be equivalent to `scale_fill_camp_rio()`. If you specify a year without branding guidelines available, then an error message will appear.

### Option 3: use individual colors from each palette

If you needed to use an additional color from a different palette you currently use, you can always call each color independently:

```{r use_individual_colors}
my_palette <- c(
  "4" = idea_colors_2019$magenta,
  "6" = idea_colors_2024$darkblue,
  "8" = camp_rio_colors_2022$forestgreen
)

cars +
  scale_fill_manual(values = my_palette)
```

Of course, always choose colors that produce a coherent visualization, since randomly selecting a palette may not create an aesthetically pleasing graph.
